### 2016-02-11 stitched.tex
ENH:  More edits.

\usepackage{booktabs}
\usepackage{changes}
\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
\linespread{1.5}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\setlist{nolistsep}
\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\setlength{\parskip}{3pt}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Nick Tustison}, color=red]{nt}
\definechangesauthor[name={James Stone}, color=magenta]{js}
\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
\definechangesauthor[name={Andy Mayer}, color=green]{am}
\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
\definechangesauthor[name={Brian Taylor}, color=purple]{bt}

@@

The random forests framework {[}18{]} is a popular machine learning
technique that has demonstrated significant utitility for supervised
segmentation tasks (e.g., normal human brain segmentation {[}19{]}) and
other computer vision applications (e.g., human gait detection
{[}20{]}). In the context of neuropathology, random forest-based
paradigms have been employed in the delineation of multiple sclerosis
lesions {[}21{]}, stroke lesions {[}22{]}, and brain tumors
{[}23--26{]}. Of note, these latter random forest approaches for brain
tumor segmentation have performed well in recent international
competitions. In response to the lack of objective comparisons between
segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS)
challenge was initiated in 2012 {[}27{]} and has continued every year
since under the auspices of the MICCAI conference.

Random forests are conceptually simple {[}18{]}. They consist of
ensembles of decision trees that are built from training data. Once
constructed, data to be classified is ``pushed'' through each decision
tree resulting in a single classification ``vote'' per tree. These votes
are then used for regression or classification of the data. Although
decision trees had been extensively studied previously, the success of
employing collections of such weak learners for boosting machine
learning performance (e.g., AdaBoost {[}28, 29{]}) influenced the
similarly sytled conglomeration of decision trees into ``forests'' with
randomized node optimization {[}30, 31{]}. Finally, Breiman {[}18{]}
improved accuracy by random sampling of the training data (i.e.,
``bagging'') resulting in the current random forest technique.

In this work, we develop a concatenated random forest framework with a
tailored contextual feature image set (both spatial and intensity-based)
for segmenting white matter hyperintensities in traumatic brain injury
cohorts. Additionally, the entire framework is provided publicly through
the well-known open-source ANTs\footnote{\url{https://github.com/stnava/ANTs}}
and ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
Further motivating the research of this work is the public availability
of the imaging data thus permitting full reproducibility of the results
reported and discussed.

@@

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions. The FLAIR and T2 images are rigidly
registered to the T1 image using the open-source Advanced Normalization
Tools (ANTs) {[}33{]}. The aligned images are then preprocessed using
the denoising algorithm of {[}35{]} followed by N4 bias correction
{[}34{]} which are then normalized to the intensity range \([0,1]\).
Although we could have used an alternative intensity standardization
algorithm (e.g., {[}37{]}), we found that a simple linear rescaling
produced better results.

@@

The T1 image is then processed via the ANTs brain extraction and normal
tissue segmentation (i.e., the pathology is discounted which is a fairly
safe assumption for the T1 image) protocols as part of the cortical
thickness pipeline described in {[}36{]}.
\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
brain has been extracted, we apply a Bayesian-based segmentation algorithm
with a template-based prior probability strategy to segment the
parenchymal tissue types.  The result is} a mask delineating the brain
parenchyma and probabilistic estimates of the cerebrospinal fluid (csf),
gray matter, white matter, deep gray matter, brain stem, and cerebellum.
These provide the expertly annotated labels for the first six tissue
labels given above. The white matter hyperintensities were manually
identified by one of the authors (J. R. S.) using the ITK-SNAP tool
{[}38{]}. Segmentation is performed using the ANTs Atropos tool {[}39{]}
and multi-model optimal symmetric shape/intensity templates {[}26{]}
created from the public MMRR data set {[}40{]} (cf Figure 3).

@@

To model the intensity information the first set of images simply
includes the preprocessed and normalized intensity FLAIR, T1, and T2
image voxel values. We also calculate a set of neighborhood statistics
(mean, standard deviation, and skewness) feature images using a
\added[id=am]{Manhattan} radius of one voxel. For each of the normalized
images, we calculate the difference in intensities with the
corresponding warped template component. Previous success in the
international brain tumor segmentation competition {[}27{]} was based on
an important set of intensity features that were created from
multi-modal templates mentioned previously {[}26{]}. We employ the same
strategy here. For example, the template difference feature image for
the FLAIR image, \(S_{FLAIR}\) is calculated as:

@@

where \(\phi_b: S \leftrightarrow \underset{b}{\leftrightsquigarrow} T\)
is the transform which maps from the individual subject space to the
template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
to take advantage of the
\added[id=hl,remark=NT:  we are using the "symmetry" of the brain in an analogous sense that if we know the left eye location and color then it is highly probabilistic that we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
bilateral symmetry of the normal brain (in terms of both shape and
intensity), and the fact that
\replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature. For the FLAIR component,
this contralateral difference image is calculated from

@@

The segmentation probability images described above are used as feature
images to provide a spatial context for the random forest model
prediction step. Additional spatial contextual feature images include
the distance maps {[}41{]} based on the csf, gray matter, and deep gray
matter images. These latter images are intended to help distinguish
white matter hyperintensities from false positives induced by the
partial voluming at the gray/white matter interface. A third set of
images are based on the voxel location within the space of the template.
The T1 image of the subject is registered to the T1 template component
using a B-spline variant {[}42{]} of the well-known ANTs Symmetric
Normalization (SyN) algorithm {[}43{]}. Since the inverse transform is
also derived as part of the registration process, we can warp the voxel
index locations back to the space of the individual subject. Note that
this is similar in motivation to the work of {[}44{]}. However, this
previous work lacks the normalization to the standard coordinate system
provided by the template to dramatically improve spatial specificity
across all subjects.

@@

As pointed out in a recent comprehensive multiple sclerosis lesion
segmentation review {[}45{]}, although the number of algorithms reported
in the literature is quite extensive, there were only four publicly
available segmentation algorithms at the time of writing of which none
are based on supervised learning. As we did for our brain tumor
segmentation algorithm {[}26{]}, all of the code described in this work
is publicly available through the open-source ANTs/ANTsR toolkits.
Through ANTsR (an add-on toolkit which, in part, bridges ANTs and the R
statistical project) we use the \texttt{randomForest} package {[}46{]}
using the default settings with 2000 trees per model and 500 randomly
selected samples per label per image.
\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}

In addition, similar to our previous offering,\footnote{\url{https://github.com/ntustison/ANTsAndArboles}}
we plan on creating a self-encapsulated example to showcase the proposed
methodology. The fact that the data will also be made available through
the FITBIR repository along with the manual labelings will facilitate
reproducibility on the part of the reader as well as any interest in
extending the proposed framework to other data sets.

@@

After performing the leave-one-out evaluation described at the end of
the previous section, we calculated the \texttt{MeanDecreaseAccuracy}
feature values for each of the 24 subjects \(\times\) 2 models per
subject \(=48\) total models. This measure (per feature, per model) is
calculated during the out-of-bag phase of the random forest model
construction and quantifies the decrease in prediction accuracy from
omitting the specified feature. In other words, this quantity helps
determine the importance of a particular feature and, although we save
such efforts for future work, this information provides us with guidance
for future feature pruning and/or additions.

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/llvAndDice.png}
\caption{Comparison with manual delineation of white matter
hyperintensities. On the left are the calculate Dice values over all
white matter hyperintensities. Note the improvement in the Dice metric
from the employment of the Stage 2 component of the processing pipeline.
(Right) Similar results can be seen by comparing the total lesion load
volume between manual and automated detection strategies. Although some
outliers are found after the Stage 2 processing in a couple subjects,
the number of outliers caused by false positives is significantly with
the second stage processing.}
\end{figure}

@@

A recent meta-analysis by Kloppenborg et al. {[}10{]} of 23
cross-sectional studies reporting MRI and concurrent neuropsychological
results in patients with heterogeneous diagnoses but without previously
diagnosed cognitive impairment, found that WMHs were associated with
cognitive deficit (effect size of -0.10, 95\% CI: -0.13 to -0.08) after
controlling for age. These studies also differed in the metric used to
measure the WMHs, including volume, \% of total intracranial volume, and
a visual rating score. The effect size for the association with
cognitive deficit in these cross-sectional studies did not differ
significantly across various cognitive domains or the method of
measuring lesion volume. Among eight longitudinal studies analyzed by
Kloppenborg et al that included a follow-up MRI and also controlled for
age, the effect size for the association of progression in WMHs and
cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
association was stronger for attention and executive function than for
memory and processing speed. Although baseline WMHs were predictive of
cognitive deficit at follow-up in the seven studies which did not repeat
MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
in the longitudinal studies that calculated progression in WMHs. In
summary, progression of WMHs seen on repeat MRI has a stronger relation
to cognitive deficit than concurrent imaging findings. These
meta-analyses support the rationale for repeating an MRI in patients
younger than 50 years whose initial scan shows WMHs. Despite the
above-described associations between WMHs, cognitive decline, increased
risk of developing dementia, and mortality, these lesions receive little
attention in current clinical workflows. When reported in a standard
neuroradiologist interpretation, they are typically handled as
incidental findings and are assigned little clinical significance. This
likely reflects the impracticality of performing a detailed assessment
of number, volume, and distribution within a qualitative
neuroradiologist interpretation as well as the lack of correlative
information on how the presence and distribution of these lesions may
inform a diagnosis and prognosis in the appropriate clinical setting. To
date, automated or semi-automated tools for the detection of WMHs have
lacked the specificity and efficiency for the mining of large-scale
datasets to generate highly granular data on whether these lesions
possess any true diagnostic or prognostic value in the setting of a
specific disease process. The present communication describes a
supervised statistical learning tool that is appropriate for the
application to such large-scale datasets.

The currently described tool is just one example of how ``deep
learning'' algorithms might be applied to aid in the diagnosis of TBI
and other disease processes through the specific identification of
features predictive of a given disease state. It is an important
demonstration of the potential power of these analytical approaches in
the rapid but comprehensive mining of information from neuroimaging
examinations. Deep learning algorithms are presently employed across a
wide variety of settings for the rapid identification of predictive
imaging features. Automobile manufacturers utilize these types of
approaches to equip self-driving vehicles to recognize and respond to
unique external surroundings through the identification of visual
information sufficiently similar to previously assimilated training
data. Similarly, in the context of the neuroimaging assessments, deep
learning approaches may allow for the rapid identification of
information predictive of disease state in an individual patient. These
approaches may be applied to the identification of macroscopically
visible lesions as described in the current communication. Additionally,
these approaches might be applied to the interrogation of imaging data
in the individual patient with a primary quantitative output metrics to
include sequences such as diffusion tensor imaging (DTI) and its
variants, functional connectivity, perfusion weighted imaging, and
cortical thickness assessments. At present, these sequences are confined
to cohort-based research studies due to the lack of available analytical
tools to assess the information in the setting of the individual
patient. Application of deep learning approaches in the context of data
with primary quantitative outputs will require large scale normative and
disease specific databases. Acquisition of these data is resource
intensive and requires a multi-center approach with harmonized scanner
between sites and correlative non-imaging clinical data. Large scale TBI
data is becoming increasingly available through activities such as the
Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research
and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European
Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of
Defense Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and
other data being consolidated through the Federal Interagency TBI
repository (FITBIR). In concert with any available high quality
normative neuroimaging data, deep learning algorithms may be well
positioned to help transform how neuroimaging is interpreted for the
clinical management of patients with this disease process.

\listofchanges

@@

>>>>

\usepackage{booktabs}
\usepackage[deletedmarkup=xout]{changes}
\usepackage[font={small},labelfont=bf,labelsep=colon]{caption}
\linespread{1.5}
\usepackage[compact]{titlesec}
\usepackage{enumitem}
\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
\setlist{nolistsep}
\titlespacing{\section}{2pt}{*0}{*0}
\titlespacing{\subsection}{2pt}{*0}{*0}
\titlespacing{\subsubsection}{2pt}{*0}{*0}
\setlength{\parskip}{3pt}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Nick Tustison}, color=red]{nt}
\definechangesauthor[name={James Stone}, color=magenta]{js}
\definechangesauthor[name={Lisa Wilde}, color=cyan]{lw}
\definechangesauthor[name={Andy Mayer}, color=green]{am}
\definechangesauthor[name={Harvey Levin}, color=blue]{hl}
\definechangesauthor[name={Brian Taylor}, color=purple]{bt}
\definechangesauthor[name={David Tate}, color=orange]{dt}

@@

The random forests framework {[}18{]} is a popular machine learning
technique that has demonstrated significant
\replaced[id=dt]{utility}{utitility} for supervised segmentation tasks
(e.g., normal human brain segmentation {[}19{]}) and other computer
vision applications (e.g., human gait detection {[}20{]}). In the
context of neuropathology, random forest-based paradigms have been
employed in the delineation of multiple sclerosis lesions {[}21{]},
stroke lesions {[}22{]}, and brain tumors {[}23--26{]}. Of note, these
latter random forest approaches for brain tumor segmentation have
performed well in recent international competitions
\replaced[id=dt]{established in response to the lack of objective comparisons between segmentation algorithms (i.e.,  the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012}{
In response to the lack of objective comparisons between segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in 2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI conference.}{[}27{]}.

Random forests are conceptually simple {[}18{]}. They consist of
ensembles of decision trees that are built from training data. Once
constructed, data to be classified is ``pushed'' through each decision
tree resulting in a single classification ``vote'' per tree. These votes
are then used for regression or classification of the data. Although
decision trees had been extensively studied \deleted[id=dt]{previously},
the success of employing collections of such weak learners for boosting
machine learning performance (e.g., AdaBoost {[}28, 29{]}) influenced
the similarly sytled conglomeration of decision trees into ``forests''
with randomized node optimization {[}30, 31{]}. Finally, Breiman
{[}18{]} improved accuracy by random sampling of \deleted[id=dt]{the}
training data (i.e., ``bagging'') resulting in the current random forest
technique \added[id=dt]{applied here}.

In this work, we develop a concatenated random forest framework with a
tailored contextual feature image set (both spatial and intensity-based)
for segmenting \replaced[id=dt]{WMHs}{white matter hyperintensities} in
\replaced[id=dt]{a arge TBI cohort}{traumatic brain injury cohorts}.
Additionally, the entire framework is provided publicly through the
well-known open-source \added[id=lw]{Advanced Normalization Tools}
(ANTs)\footnote{\url{https://github.com/stnava/ANTs}} and
ANTsR\footnote{\url{https://github.com/stnava/ANTsR}} toolkits.
\replaced[id=dt]{Further motivating this research is the availability of several large publicly available  imaging data sets that permits testing reproducibility of this automate routine for WMH segmentation and quantification.}{Further motivating this research of this work is
the public availability of the imaging data thus permitting full reproducibility
of the results reported and discussed.}

@@

As mentioned previously, input for each subject comprises FLAIR, T1-,
and T2-weighted acquisitions. The FLAIR and T2 images are rigidly
registered to the T1 image using the open-source Advanced Normalization
Tools (ANTs) {[}33{]}. The aligned images are then preprocessed using
the denoising algorithm of {[}35{]} followed by N4 bias correction
{[}34{]} which are then normalized to the intensity range \([0,1]\).
Although we could have used an alternative intensity standardization
algorithm (e.g., {[}37{]}), we found that a simple linear rescaling
produced better results similar \added[id=dt]{to previous work}{[}26{]}.

@@

The T1 image is then processed via the ANTs brain extraction and normal
tissue segmentation (i.e., the pathology is discounted which is a fairly
safe assumption for the T1 image) protocols as part of the cortical
thickness pipeline described in {[}36{]}.
\added[id=hl]{This protocol involves preprocessing using N4 bias correction followed by a template-based strategy for brain extraction.  Once the
brain has been extracted, we apply a Bayesian-based segmentation algorithm
with a template-based prior probability strategy to segment the
parenchymal tissue types.  The result is} a mask delineating the brain
parenchyma and probabilistic estimates of the
\replaced[id=dt]{CSF}{cerebrospinal fluid (csf)}, gray matter, white
matter, deep gray matter, brain stem, and cerebellum. These provide the
expertly annotated labels for the first six tissue labels given above.
The \replaced[id=dt]{WMHs}{white matter hyperintensities} were manually
identified by one of the authors (J. R. S.) using the ITK-SNAP tool
{[}38{]}. Segmentation is performed using the ANTs Atropos tool {[}39{]}
and multi-model optimal symmetric shape/intensity templates {[}26{]}
created from the public MMRR data set {[}40{]} (cf Figure 3).

@@

To model the intensity information the first set of images simply
includes the preprocessed and normalized intensity FLAIR, T1, and T2
image voxel values. We also calculate a set of neighborhood statistics
(mean, standard deviation, and skewness) feature images using a
\added[id=am]{Manhattan} radius of one voxel
\added[id=nt]{given the typical size
of individual WMHs}. For each of the normalized images, we calculate the
difference in intensities with the corresponding warped template
component. Previous success in the international brain tumor
segmentation competition {[}27{]} was based on an important set of
intensity features that were created from multi-modal templates
mentioned previously {[}26{]} \added[id=nt]{and listed in Table 1}. We
employ the same strategy here. For example, the template difference
feature image for the FLAIR image, \(S_{FLAIR}\) is calculated as:

@@

where \(\phi_b: S \leftrightarrow \underset{b}{\leftrightsquigarrow} T\)
is the transform which maps from the individual subject space to the
template space and \(T_{FLAIR}\) is the FLAIR template component. Also,
to take advantage of the
\added[id=hl,remark=NT:  we are using the gross bilateral symmetry of the brain in an analogous sense that if we know the left eye location and color then we can accurately predict the location and color of the right eye even though a particular face might not be exactly symmetric]{gross}
bilateral symmetry of the normal brain (in terms of both shape and
intensity), and the fact that
\replaced[id=bt]{WMHs do not generally manifest symmetrically across hemispheres}{the presence of WMHs violates that assumption},
we use the symmetric templates to compute the contralateral intensity
differences as an additional intensity feature. For the FLAIR component,
this contralateral difference image is calculated from

@@

The segmentation probability images described above are used as feature
images to provide a spatial context for the random forest model
prediction step. Additional spatial contextual feature images include
the distance maps {[}41{]} based on the csf, gray matter, and deep gray
matter images. These latter images are intended to help distinguish
white matter hyperintensities from false positives induced by the
partial voluming at the gray/white matter interface. A third set of
images are based on the voxel location within the space of the template.
The T1 image of the subject is registered to the T1 template component
using a B-spline variant {[}42{]} of the well-known ANTs Symmetric
Normalization (SyN) algorithm {[}43{]}. Since the inverse transform is
also derived as part of the registration process, we can warp the voxel
index locations back to the space of the individual subject
\replaced[id=nt]{which motivates similar work by others}{
Note that this is similar in motivation to the work
of} {[}44{]}. However, this previous work lacks the normalization to the
standard coordinate system provided by the template to dramatically
improve spatial specificity across all subjects.

@@

As pointed out in a recent comprehensive multiple sclerosis lesion
segmentation review {[}45{]}, although the number of algorithms reported
in the literature is quite extensive, there were only four publicly
available segmentation algorithms at the time of writing
\added[id=am]{this article}.
\replaced[id=dt]{In contrast to the current work,}{of which} none are
based on supervised learning. As we did for our brain tumor segmentation
algorithm {[}26{]}, all of the code described in this work is publicly
available through the open-source ANTs/ANTsR toolkits. Through ANTsR (an
add-on toolkit which, in part, bridges ANTs and the R statistical
project) we use the \texttt{randomForest} package {[}46{]} using the
default settings with 2000 trees per model and 500 randomly selected
samples per label per image.
\added[id=bt]{Note that we saw little variation in performance when these parameters were changed which is consistent with our previous experience.}

In addition, similar to our previous offering,\footnote{\url{https://github.com/ntustison/ANTsAndArboles}}
we plan on creating a self-encapsulated example to showcase the proposed
methodology. The fact that the data will also be made available through
the
\replaced[id=lw]{Federal Interagency Traumatic Brain Injury Research (FITBIR)}{FITBIR}
repository along with the manual labelings will facilitate
reproducibility on the part of the reader as well as any interest in
extending the proposed framework to other data sets.

@@

After performing the leave-one-out
evaluation\deleted[id=dt]{described at the end of the previous section},
we calculated the \texttt{MeanDecreaseAccuracy} feature values for each
of the 24 subjects \(\times\) 2 models per subject \(=48\) total models.
This measure (per feature, per model) is calculated during the
out-of-bag phase of the random forest model construction and quantifies
the decrease in prediction accuracy from omitting the specified feature.
In other words, this quantity helps determine the importance of a
particular feature and, although we save such efforts for future work,
this information provides us with guidance for future feature pruning
and/or additions.

@@

\begin{figure}[htbp]
\centering
\includegraphics{Figures/llvAndDice.png}
\caption{\added[id=am]{Voxelwise} comparison with manual delineation of
white matter hyperintensities. On the left are the
calculate\added[id=am]{d} Dice values over all white matter
hyperintensities. Note the improvement in the Dice metric from the
employment of the Stage 2 component of the processing pipeline. (Right)
Similar results can be seen by comparing the total lesion load volume
between manual and automated detection strategies. Although some
outliers are found after the Stage 2 processing in a couple subjects,
the number of outliers caused by false positives is
\added[id=lw]{decreased} significantly with the second stage
processing.}
\end{figure}

@@

\replaced[id=dt]{The more}{A} recent meta-analysis by Kloppenborg et al.
{[}10{]} of 23 cross-sectional studies reporting MRI and concurrent
neuropsychological results in patients with heterogeneous diagnoses but
without previously diagnosed cognitive impairment, found that WMHs were
associated with cognitive deficit (effect size of -0.10, 95\% CI: -0.13
to -0.08) after controlling for age. These studies also differed in the
metric used to measure the WMHs, including volume, \% of total
intracranial volume, and a visual rating score. The effect size for the
association with cognitive deficit in these cross-sectional studies did
not differ significantly across various cognitive domains or the method
of measuring lesion volume. Among eight longitudinal studies analyzed by
Kloppenborg et al that included a follow-up MRI and also controlled for
age, the effect size for the association of progression in WMHs and
cognitive impairment was -0.16 (95\% CI:-0.27 to -0.09). This
association was stronger for attention and executive function than for
memory and processing speed. Although baseline WMHs were predictive of
cognitive deficit at follow-up in the seven studies which did not repeat
MRI, the effect size was smaller {[}-0.10 (95\% CI: 0.13 to -0.05) than
in the longitudinal studies that calculated progression in WMHs. In
summary, progression of WMHs seen on repeat MRI has a stronger relation
to cognitive deficit than concurrent imaging findings. These
meta-analyses support the rationale for repeating an MRI in patients
younger than 50 years whose initial scan shows WMHs. Despite the
above-described associations between WMHs, cognitive decline, increased
risk of developing dementia, and mortality, these lesions receive little
attention in current clinical workflows. When reported in a standard
neuroradiologist interpretation, they are typically handled as
incidental findings and are assigned little clinical significance. This
likely reflects the impracticality of performing a detailed assessment
of number, volume, and distribution within a qualitative
neuroradiologist interpretation as well as the lack of correlative
information on how the presence and distribution of these lesions may
inform a diagnosis and prognosis in the appropriate clinical setting. To
date, automated or semi-automated tools for the detection of WMHs have
lacked the specificity and efficiency for the mining of large-scale
datasets to generate highly granular data on whether these lesions
possess any true diagnostic or prognostic value in the setting of a
specific disease process. The present communication describes a
supervised statistical learning tool that is appropriate for the
application to such large-scale datasets.

The currently described tool is just one example of how ``supervised
learning'' algorithms might be applied to aid in the diagnosis of TBI
and other disease processes through the specific identification of
features predictive of a given disease state. It is an important
demonstration of the potential power of these analytical approaches in
the rapid but comprehensive mining of information from neuroimaging
examinations. Supervised learning algorithms are presently employed
across a wide variety of settings for the rapid identification of
predictive imaging features {[}50--53{]}. Automobile manufacturers
utilize these types of approaches to equip self-driving vehicles to
recognize and respond to unique external surroundings through the
identification of visual information sufficiently similar to previously
assimilated training data {[}54, 55{]}. Similarly, in the context of the
neuroimaging assessments, deep learning approaches may allow for the
rapid identification of information predictive of disease state in an
individual patient. These approaches have been applied to the
segmentation of macroscopically visible structures {[}50--53{]}.
Additionally, these approaches might be applied to the interrogation of
imaging data in the individual patient with a primary quantitative
output metrics to include sequences such as diffusion tensor imaging
(DTI) and its variants, functional connectivity, perfusion weighted
imaging, and cortical thickness assessments. At present, these advanced
neuroimaging sequences are confined to cohort-based research studies due
to the lack of available analytical tools to assess the information in
the setting of the individual patient {[}56{]}. Application of deep
learning approaches in the context of data with primary quantitative
outputs will require large scale normative and disease specific
databases. Building these large scale imaging libraries is resource
intensive and requires a multi-center approach with harmonized scanners
between sites and correlative non-imaging clinical data. Large scale TBI
data is becoming increasingly available through activities such as the
Chronic Effects of Neurotrauma Consortium (CENC), Transforming Research
and Clinical Knowledge in TBI (TRACK-TBI), Collaborative European
Neurotrauma Effectiveness Research in TBI (CENTER-TBI), Department of
Defense Alzheimer's Disease Neuroimaging Initiative (DOD-ADNI), and
other data being consolidated through FITBIR. In concert with any
available high quality normative neuroimaging data, deep learning
algorithms may be well positioned to help transform how neuroimaging is
interpreted for the clinical management of patients with this disease
process.

@@

\hypertarget{ref-Plis:2014aa}{}
50. Plis, S. M., Hjelm, D. R., Salakhutdinov, R., Allen, E. A.,
Bockholt, H. J., Long, J. D., Johnson, H. J., Paulsen, J. S., Turner, J.
A., and Calhoun, V. D. ``\textbf{Deep Learning for Neuroimaging: A
Validation Study}'' \emph{Front Neurosci} 8, (2014): 229.
doi:\href{https://doi.org/10.3389/fnins.2014.00229}{10.3389/fnins.2014.00229}

\hypertarget{ref-Suk:2015aa}{}
51. Suk, H.-I., Lee, S.-W., Shen, D., and Alzheimer's Disease
Neuroimaging Initiative. ``\textbf{Deep Sparse Multi-Task Learning for
Feature Selection in Alzheimer's Disease Diagnosis}'' \emph{Brain Struct
Funct} (2015):
doi:\href{https://doi.org/10.1007/s00429-015-1059-y}{10.1007/s00429-015-1059-y}

\hypertarget{ref-Li:2014aa}{}
52. Li, R., Zhang, W., Suk, H.-I., Wang, L., Li, J., Shen, D., and Ji,
S. ``\textbf{Deep Learning Based Imaging Data Completion for Improved
Brain Disease Diagnosis}'' \emph{Med Image Comput Comput Assist Interv}
17, no. Pt 3 (2014): 305--12.

\hypertarget{ref-Liu:2015aa}{}
53. Liu, S., Liu, S., Cai, W., Che, H., Pujol, S., Kikinis, R., Feng,
D., Fulham, M. J., and ADNI. ``\textbf{Multimodal Neuroimaging Feature
Learning for Multiclass Diagnosis of Alzheimer's Disease}'' \emph{IEEE
Trans Biomed Eng} 62, no. 4 (2015): 1132--40.
doi:\href{https://doi.org/10.1109/TBME.2014.2372011}{10.1109/TBME.2014.2372011}

\hypertarget{ref-HadsellSBESKML09}{}
54. Hadsell, R., Sermanet, P., Ben, J., Erkan, A., Scoffier, M.,
Kavukcuoglu, K., Muller, U., and LeCun, Y. ``\textbf{Learning Long-Range
Vision for Autonomous Off-Road Driving}'' \emph{J. Field Robotics} 26,
no. 2 (2009): 120--144.
doi:\href{https://doi.org/10.1002/rob.20276}{10.1002/rob.20276},
Available at \url{http://dx.doi.org/10.1002/rob.20276}

\hypertarget{ref-FarabetCNL12}{}
55. Farabet, C., Couprie, C., Najman, L., and LeCun, Y. ``\textbf{Scene
Parsing with Multiscale Feature Learning, Purity Trees, and Optimal
Covers}'' \emph{Proceedings of the 29th international conference on
machine learning, ICML 2012, edinburgh, scotland, uK, june 26 - july 1,
2012} (2012): Available at \url{http://icml.cc/discuss/2012/319.html}

\hypertarget{ref-Mayer:2014aa}{}
56. Mayer, A. R., Bedrick, E. J., Ling, J. M., Toulouse, T., and Dodd,
A. ``\textbf{Methods for Identifying Subject-Specific Abnormalities in
Neuroimaging Data}'' \emph{Hum Brain Mapp} 35, no. 11 (2014): 5457--70.
doi:\href{https://doi.org/10.1002/hbm.22563}{10.1002/hbm.22563}
